---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
```{r}

data <- read.csv(file = "activity.csv",header = TRUE);
#remove uncomplete data
compData <- data[complete.cases(data),]

#convert date column
library(lubridate)
compData[,2] <- ymd(compData[,2])
data[,2] <- ymd(data[,2])

```


## What is mean total number of steps taken per day?

* Make a histogram of the total number of steps taken each day

```{r}
suppressPackageStartupMessages(library(dplyr))
group_step_by_day <- compData %>% 
                        group_by(date) %>% 
                        summarise( total = sum(steps) )
 

library(ggplot2)
g1 <- ggplot(
        group_step_by_day, 
        aes(date ,total  )
        )

g1 + geom_histogram( 
        aes(fill = total),
        binwidth = 0.1,stat="identity" 
        ) +
xlab("Date")  + 
ylab("Total Steps") 

 
```


* Calculate and report the mean and median total number of steps taken per day

```{r results='asis'}
library(DT)
mean_median_step_by_day <- compData %>% 
                        group_by(date) %>% 
                        summarise( 
                                Mean = mean(steps), 
                                Median = median(steps) 
                                )


mean_median_step_by_day$date <- as.factor(mean_median_step_by_day$date)
datatable(mean_median_step_by_day,options = list( autoWidth = TRUE ,pageLength = 25 ))

```


## What is the average daily activity pattern?
```{r }

average_dayly <- compData %>% 
                        group_by(date) %>% 
                        group_by(interval) %>% 
                        summarise( 
                                Mean = mean(steps)
                                )


max_interval <- average_dayly[average_dayly$Mean==max(average_dayly$Mean),]$interval;
 
g <- ggplot(
        average_dayly, 
        aes(interval,Mean)
        )

g + geom_line() + geom_vline(xintercept = max_interval  , colour="green", linetype = "longdash")+
         annotate("text", x = max_interval, y = -2, label = max_interval)

```

The 5-minute interval that contains the maximum number of steps on average across all the days in the dataset is **`r max_interval`**.


## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
number_NA <- sum(!complete.cases(data))
```
  
The total number of rows with NAs is **`r number_NA`**.



Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
  
nas <- data[!complete.cases(data), ]

for(i in 1:dim(nas)[1]){
        nas[i,]$steps <- average_dayly[average_dayly$interval==nas[i,]$interval,]$Mean        
}


#Create a new dataset that is equal to the original dataset but with the missing data filled in.
data_full <- data
data_full[!complete.cases(data_full), ] <- nas

 


full_group_step_by_day <- data_full %>% 
                        group_by(date) %>% 
                        summarise( total = sum(steps) )
 

#Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

h <- ggplot(
        full_group_step_by_day, 
        aes(date ,total  )
        )

h + geom_histogram( 
        aes(fill = total),
        binwidth = 0.1,stat="identity" 
        ) +
xlab("Date")  + 
ylab("Total Steps")  




full_mean_median_step_by_day <- data_full %>% 
                        group_by(date) %>% 
                        summarise( 
                                Mean = mean(steps), 
                                Median = median(steps) 
                                )


full_mean_median_step_by_day$date <- as.factor(full_mean_median_step_by_day$date)
datatable(full_mean_median_step_by_day,options = list( autoWidth = TRUE ,pageLength = 25 ))
```

###Do these values differ from the estimates from the first part of the assignment?   
Yes  

```{r}
p <- ggplot() + 
        geom_histogram( data = full_group_step_by_day,
        aes(date ,total ,color = "Full"),
        binwidth = 0.1,stat="identity" 
        ) +
        geom_histogram( data = group_step_by_day,
        aes(date ,total ,color = "without NAs"),
        binwidth = 0.1,stat="identity" 
        ) +
        
xlab("Date")  + 
ylab("Total Steps")

p
```



###What is the impact of imputing missing data on the estimates of the total daily number of steps?  
increase Median of total daily number of steps  


 
## Are there differences in activity patterns between weekdays and weekends?

```{r}

#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
suppressPackageStartupMessages(library(chron))
data_full_week <- data_full %>% mutate( type = ifelse( is.weekend(date),yes = "weekend",no = "weekday" ) )


full_average_dayly_we <- data_full_week %>% 
                        filter(type=="weekend") %>% 
                        group_by(date) %>% 
                        group_by(interval) %>% 
                        summarise( 
                                Mean = mean(steps) 
                                )
full_average_dayly_we$type <-"weekend"

full_average_dayly_wd <- data_full_week %>% 
                        filter(type=="weekday") %>% 
                        group_by(date) %>% 
                        group_by(interval) %>% 
                        summarise( 
                                Mean = mean(steps) 
                                )
 
full_average_dayly_wd$type <- "weekday"

full_avg_dayly <- rbind(full_average_dayly_wd,full_average_dayly_we)


g <- ggplot(full_avg_dayly,aes(interval,Mean))

g + geom_line( )+facet_grid(type~.)

 

```

